---
title: "1: Loading AIBS Data"
author: "Sonny"
output: html_document
---

This document details the code for loading AIBS human brain data as of August 4, 2020. This is a new version of the analyses performed previously in May-Jun 2020 (see "Run_V1" folder for those details/scripts.) Data is downloaded from https://portal.brain-map.org/atlases-and-data/rnaseq/human-multiple-cortical-areas-smart-seq .

First, we load and look at the metadata to see what we have available and make our data/sample selection (if/as needed).

```{r, loading and reviewing meta-data, filtering data}

new_metadata <- read.csv("/external/rprshnas01/netdata_kcni/stlab/AIBS_scRNAseq_2019/human/metadata.csv", stringsAsFactors=FALSE) #this file corresponds to https://idk-etl-prod-download-bucket.s3.amazonaws.com/aibs_human_ctx_smart-seq/metadata.csv , we are using FALSE for strings as factors so that we can edit and order them later

names(new_metadata) #see columns that we have to work with, "outlier_call" and "subclass_label" are of interest to us here

table(new_metadata$outlier_call, useNA = "ifany") #see the outlier data; there are 1985 "outliers", likely corresponding to "tome_sample_meta_filtered$class_label == 'Exclude'" from Run_V1

new_metadata_filtered <- new_metadata[new_metadata$outlier_call == "False",] #filter out outliers

table(new_metadata_filtered$subclass_label, useNA = "ifany") #for this demo/workbook, we will be working with subclass labels; this is to let us see what data we have

#subclass_exclusion_list <- c("IT", "L4 IT", "L5 ET", "L6b") #cell subclasses that we are excluding; in this case, we are excluding them because they are very general and vague and could steal/obscure markers from the other, more defined cell types (if needed/as desired)

#new_metadata_filtered <- new_metadata_filtered[!(new_metadata_filtered$subclass_label %in% subclass_exclusion_list),] #filter data based on subclass exclusion (if needed/as desired)

table(new_metadata_filtered[,c("subclass_label", "region_label")], useNA = "ifany") #we can see from this that non-neuronal cell types have drastically decreased sample sizes when split to brain regions; we will pick neurons from MTG and CgG as well as non-neuronal cell types from all brain regions for our demo/this project

table(new_metadata_filtered[,c("class_label", "region_label")], useNA = "ifany") #re-illustrate the point above

new_metadata_filtered <- new_metadata_filtered[!(new_metadata_filtered$class_label!="Non-neuronal") | (new_metadata_filtered$region_label %in% c("CgG", "MTG")),] #filter out neurons from regions other than CgG or MTG

table(new_metadata_filtered[,c("class_label", "region_label")], useNA = "ifany") #see our sample distribution now
table(new_metadata_filtered[,c("subclass_label", "region_label")], useNA = "ifany") #see our sample distribution now

new_metadata_filtered$NeuN <- new_metadata_filtered$class_label #make a new variable for classifying as neuron/nonneuron
new_metadata_filtered[new_metadata_filtered$NeuN != "Non-neuronal", "NeuN"] <- "Neuronal" #change gabaergic/glutamatergic to Neuronal
table(new_metadata_filtered[,c("NeuN", "region_label")], useNA = "ifany") #see our sample distribution now

new_metadata_filtered$NeuN_Region <- paste0(new_metadata_filtered$region_label, "_", new_metadata_filtered$NeuN) #this variable will help us remove MTG or CgG neurons later in the analysis
table(new_metadata_filtered$NeuN_Region, useNA = "ifany") #final view of our sample distribution

```

Now we can load our data.


```{r, loading_data}

new_count_matrix <- read.csv("/external/rprshnas01/netdata_kcni/stlab/AIBS_scRNAseq_2019/human/matrix.csv", stringsAsFactors=FALSE) #this file corresponds to https://idk-etl-prod-download-bucket.s3.amazonaws.com/aibs_human_ctx_smart-seq/matrix.csv 

new_count_matrix[1:10,1:10] #preview of uploaded data
row.names(new_count_matrix) <- new_count_matrix$sample_name #set rownames
new_count_matrix <- new_count_matrix[new_metadata_filtered$sample_name,] #filter for the cells we specified previously
new_count_matrix <- new_count_matrix[,2:ncol(new_count_matrix)] #the df so far has samples as the first column, which is generally not desired for many situations. Thus, we set the row names to the first column and remove the first column; careful to run this only ONCE

new_count_matrix[1:10,1:10] #preview of uploaded data again

```

The count matrix is ready for work/conversion into another workable format.

This is the end of this script, for now. We will take this data into other scripts and convert into appropriate format as desired, ultimately leading to analysis.
