---
title: "Marker evaluation and comparisons"
author: "Sonny"
output: html_notebook
---

This notebook contains code and work evaluating marker gene lists, both internally and against other lists. For the example in this notebook, we are evaluating the reults from "Seurat_Implementation.Rmd".  

First, we can do a quick check that no genes were determined to be markers for more than one group. The fastest way to do this is to check that all "gene" values are unique.

```{r, unique marker check}

# if there are multiple iterations/versions of results, like in the later parts of Seurat_Implementation.Rmd, we can select which iteration to use and/or filter for comparisons (instead of Seu_AIBS_markers)

length(unique(Seu_AIBS_markers$gene)) #this shows us that the number of unique gene entries are different compared to the number of rows, indicating duplicates
duplicated_markers <- Seu_AIBS_markers[duplicated(Seu_AIBS_markers$gene),"gene"] #get list of duplicated markers
Seu_AIBS_markers[Seu_AIBS_markers$gene %in% duplicated_markers,] #see the rows with duplicated markers
Seu_AIBS_markers_filtered <- Seu_AIBS_markers[!(Seu_AIBS_markers$gene %in% duplicated_markers),] #remove duplicated markers

# if using MTG results, for example
duplicated_markers <- Seu_AIBS_markers_MTG[duplicated(Seu_AIBS_markers_MTG$gene),"gene"] #get list of duplicated markers
Seu_AIBS_markers_filtered <- Seu_AIBS_markers_MTG[!(Seu_AIBS_markers_MTG$gene %in% duplicated_markers),] #remove duplicated markers

```

Now we want to compare our results to previous/other marker gene lists. To do this, we load the data and convert it to a dataframe comparable to our Seurat results.

```{r, comparing lists}

human_markers_quick_sel <- readRDS("~/git/MarkerSelection/Data/human_markers_quick_sel.rds") #loading a previous marker gene list we've been working with in our lab

human_markers_quick_sel #see the previous list of markers

# we want to convert this list into a dataframe comparable with our Seurat results

human_markers_quick_sel_df <- Seu_AIBS_markers_filtered[0,c("cluster", "gene")] #initializing a blank df, based on Seurat result df [1]

for (cluster_name in names(human_markers_quick_sel)) { #for each cell grouping
  
  temp_df <- as.data.frame(human_markers_quick_sel[cluster_name]) #convert list to a dataframe
  temp_df$cluster <-cluster_name #add cell grouping/cluster/etc. as a second column of values
  temp_df <- temp_df[,c(2,1)] #reverse order of columns
  colnames(temp_df)[2] <- "gene" #name the column of marker genes
  
  #now we can add this structure-converted list of gene markers to our df initialized in [1] (line 19)
  human_markers_quick_sel_df <- rbind(human_markers_quick_sel_df, temp_df) 
  
  }

remove(temp_df) #we can get rid of the temporary df

head(human_markers_quick_sel_df) #we can now see the converted list of gene markers in df format

```

Now we just need to convert the "cluster"/grouping names to match between our reference list and our results/Seurat list.
For now, we do this manually due to the complexity and differences between the two lists. 
The reference names can be converted to the result names or vice-versa, we do the latter because it has more cell groups and multiple cell groups from it can compare to one in the reference cell-group.'

```{r, converting grouping variable}

Seu_AIBS_markers_filtered$adapted_cluster_name <- Seu_AIBS_markers_filtered$cluster #copy the cluster column to start a new column of adapted cluster names

unique(human_markers_quick_sel_df$cluster) #these are the reference names that we want to adapt/convert to
unique(Seu_AIBS_markers_filtered$cluster) #these are the names that we want to change to match the above

# manual replacements:

Seu_AIBS_markers_filtered$adapted_cluster_name <- replace(as.character(Seu_AIBS_markers_filtered$adapted_cluster_name), Seu_AIBS_markers_filtered$adapted_cluster_name == "Astro_FGFR3", "Astrocyte")

Seu_AIBS_markers_filtered$adapted_cluster_name <- replace(as.character(Seu_AIBS_markers_filtered$adapted_cluster_name), Seu_AIBS_markers_filtered$adapted_cluster_name == "Endo_CLDN5", "Endothelial")

Seu_AIBS_markers_filtered$adapted_cluster_name <- replace(as.character(Seu_AIBS_markers_filtered$adapted_cluster_name), Seu_AIBS_markers_filtered$adapted_cluster_name == "Micro_C1QC", "Microglia")

Seu_AIBS_markers_filtered$adapted_cluster_name <- replace(as.character(Seu_AIBS_markers_filtered$adapted_cluster_name), Seu_AIBS_markers_filtered$adapted_cluster_name == "OPC_MYT1", "OPC")

Seu_AIBS_markers_filtered$adapted_cluster_name <- replace(as.character(Seu_AIBS_markers_filtered$adapted_cluster_name), Seu_AIBS_markers_filtered$adapted_cluster_name == "Inh_PVALB", "PVALB")

Seu_AIBS_markers_filtered$adapted_cluster_name <- replace(as.character(Seu_AIBS_markers_filtered$adapted_cluster_name), Seu_AIBS_markers_filtered$adapted_cluster_name == "Inh_SST", "SST")

Seu_AIBS_markers_filtered$adapted_cluster_name <- replace(as.character(Seu_AIBS_markers_filtered$adapted_cluster_name), Seu_AIBS_markers_filtered$adapted_cluster_name == "Inh_VIP", "VIP")

# multi-replacements (also manual):

Seu_AIBS_markers_filtered$adapted_cluster_name <- replace(as.character(Seu_AIBS_markers_filtered$adapted_cluster_name), Seu_AIBS_markers_filtered$adapted_cluster_name == "Inh_LAMP5", "LAMP5.PAX6.Other")

Seu_AIBS_markers_filtered$adapted_cluster_name <- replace(as.character(Seu_AIBS_markers_filtered$adapted_cluster_name), Seu_AIBS_markers_filtered$adapted_cluster_name == "Inh_PAX6", "LAMP5.PAX6.Other")

unique(Seu_AIBS_markers_filtered$adapted_cluster_name) #see our adapted cluster/group names

```

Now we can compare the two lists.

```{r, list comparison}

for (cluster_name in unique(Seu_AIBS_markers_filtered$adapted_cluster_name)) { #for each group (subclass, in this case)
  
  ref_num <- length(human_markers_quick_sel_df[human_markers_quick_sel_df$cluster == cluster_name, "gene"]) #number of marker genes in reference
  res_num <- length(Seu_AIBS_markers_filtered[Seu_AIBS_markers_filtered$adapted_cluster_name == cluster_name, "gene"]) #number of marker genes in results
  inter_num <- length(intersect(human_markers_quick_sel_df[human_markers_quick_sel_df$cluster == cluster_name, "gene"] ,Seu_AIBS_markers_filtered[Seu_AIBS_markers_filtered$adapted_cluster_name == cluster_name, "gene"])) #number of overlapping marker genes
  
  print(c(cluster_name, ref_num, res_num, inter_num)) #print the results
  
  print("")
  
  print(intersect(human_markers_quick_sel_df[human_markers_quick_sel_df$cluster == cluster_name, "gene"] ,Seu_AIBS_markers_filtered[Seu_AIBS_markers_filtered$adapted_cluster_name == cluster_name, "gene"])) #print list of overlapping marker genes
  
  print("")
  
  }

```

Similarly, we can compare the results from different analyses done by Seurat with our dataset (in this case, MAST vs roc or ACC vs MTG, which we performed in Seurat_Implementation) selected in the FindAllMarkers function.

```{r, multi-test comparison}

Seu_AIBS_markers_multicomp <- Seu_AIBS_markers_roc # to compare vs ACC roc
Seu_AIBS_markers_multicomp <- Seu_AIBS_markers_MTG # to compare vs MTG MAST
Seu_AIBS_markers_multicomp <- Seu_AIBS_markers_MTGroc # to compare vs MTG roc

for (cluster_name in unique(Seu_AIBS_markers_filtered$cluster)) { #for each group (subclass, in this case)
  
  ref_num <- length(unique(Seu_AIBS_markers_multicomp[Seu_AIBS_markers_multicomp$cluster == cluster_name, "gene"])) #number of marker genes in reference
  res_num <- length(Seu_AIBS_markers_filtered[Seu_AIBS_markers_filtered$cluster == cluster_name, "gene"]) #number of marker genes in results
  inter_num <- length(intersect(unique(Seu_AIBS_markers_multicomp[Seu_AIBS_markers_multicomp$cluster == cluster_name, "gene"]) ,Seu_AIBS_markers_filtered[Seu_AIBS_markers_filtered$cluster == cluster_name, "gene"])) #number of overlapping marker genes
  
  print(c(cluster_name, ref_num, res_num, inter_num)) #print the results
  
  print("")
  
  print(intersect(Seu_AIBS_markers_multicomp[Seu_AIBS_markers_multicomp$cluster == cluster_name, "gene"] ,Seu_AIBS_markers_filtered[Seu_AIBS_markers_filtered$cluster == cluster_name, "gene"])) #print list of overlapping marker genes, if desired
  
  print("")
  
  }


```

Since the matches are so close between ACC-MAST results and the ACC-roc results, we can combine the two (the the downstream user/assessor can filter/use whichever they want).

```{r, combining MAST and roc, getting output result}

# we are using the "_multicomp" df to allow for flexibility upstream to use other df instead of "_roc", such as "_MTGroc"; similarly, we are making a result DF instead of a specifically named result df for most of the code below

duplicated_markers <- Seu_AIBS_markers_multicomp[duplicated(Seu_AIBS_markers_multicomp$gene),"gene"] #get list of duplicated markers from the roc test
Seu_AIBS_markers_multicomp[Seu_AIBS_markers_multicomp$gene %in% duplicated_markers,] #see the rows with duplicated markers
roc_df <- Seu_AIBS_markers_multicomp[!(Seu_AIBS_markers_multicomp$gene %in% duplicated_markers),] #remove duplicated markers
Result_df <- merge(Seu_AIBS_markers_filtered[,c(1,5,7,8)], roc_df, by = "gene", all.x = TRUE, all.y = TRUE) #combine the marker df
Result_df <- Result_df[,c(1,10,4,8,9,6,2,3,5,7)] #reorder columns
colnames(Result_df)[c(6:10)] <- c("avg_logFC","MAST_p_val","MAST_p_val_adj", "roc_myAUC", "roc_power") #rename columns to clarify which test generated which statistic

# finally, we want to add entrez IDs to the output/result df
# our mgi-to-entrez ID conversion comes from a file from AIBS http://celltypes.brain-map.org/api/v2/well_known_file_download/694416044

human_MTG_2018_06_14_genes_rows <- read_csv("~/git/MarkerSelection/Data/Inputs/human_MTG_2018-06-14_genes-rows.csv") #loading the conversion data
Result_df <- merge(Result_df, human_MTG_2018_06_14_genes_rows[,c(1,3)], by = "gene") #add entrez ids
Result_df <- Result_df[,c(1,11,2:10)] #reorder columns

ACC_results <- Result_df #store the results
MTG_results <- Result_df #store the results

write.csv(ACC_results, "/external/rprshnas01/kcni/ychen/git/MarkerSelection/Data/Outputs/ACC_results.csv")
write.csv(MTG_results, "/external/rprshnas01/kcni/ychen/git/MarkerSelection/Data/Outputs/MTG_results.csv")

```

There are some other checks/visualizations we can do of the data as well.

We can check how correlated all the queried subclasses are to one another based on gene averages per group.

```{r, group averages and their correlations to one another}

# since our data is already in Seurat, we can use its functions for getting group averages

library(Seurat) #if needed (if it is not loaded when you need to run the code below)

Group_averages <- AverageExpression(Seu_AIBS_obj) #getting group averages (returns a list of 1 df)
Group_averages <- Group_averages$RNA #extracting the dataframe format of the results

Group_correlations <- cor(Group_averages, method = "spearman") #getting spearman coefficients of our query groups

# to visualize the confusion matrix/find most similar groups, we can plot/make a dendrogram

heatmap(Group_correlations) #one way/function to plot/make the dendrogram

library(gplots)
heatmap.2(Group_correlations) #another way to plot/make the dendrogram

```

Similarly, we can see how markers correlate to one another, visualized by a dendrogram.

```{r, marker intercorrelation}

# we can take all cells from our Seurat object

AIBS_lnCPM_InEx_ACC <- as.data.frame(Seu_AIBS_obj_ACC_it2[["RNA"]]@data) #get lnCPM matrix

# or we might want to look at correlations only within a cell group, Inh_SST cells in this case

Seu_AIBS_obj_temp <- subset(Seu_AIBS_obj, idents = "Inh_SST") #we want all cells except for these two groups
AIBS_lnCPM_InEx_ACC <- as.data.frame(Seu_AIBS_obj_temp[["RNA"]]@data) #get lnCPM matrix
remove(Seu_AIBS_obj_temp)

# whichever lnCPM matrix you select, we can subset for our genes of interest and calculate + plot the correlations

Marker_matrix <- t(AIBS_lnCPM_InEx_ACC[row.names(AIBS_lnCPM_InEx_ACC) %in% unique(ACC_results[ACC_results$cluster == "Inh_SST","gene"]),]) #subset for markers of interest

Marker_matrix <- as.matrix((Marker_matrix > 0) + 0) #if we instead want to work with a binary present/absent matrix (optional)

Marker_correlations <- cor(Marker_matrix, method = "spearman") #getting spearman coefficients of our query groups

heatmap(Marker_correlations) #one way/function to plot/make the dendrogram

```


We can also see how markers within a cell group might look on a dendrogram just based on their expression.

```{r, marker dendrogram}

Group_averages <- AverageExpression(Seu_AIBS_obj_ACC_it2) #getting group averages (returns a list of 1 df)
Group_averages <- Group_averages$RNA #extracting the dataframe format of the results

unique(ACC_results$cluster) #see options for which group we want to look at
Group_averages <- Group_averages[row.names(Group_averages) %in% unique(ACC_results[ACC_results$cluster == "Inh_SST","gene"]),] #subset for marker genes

heatmap(as.matrix(Group_averages))

```

This may be better visualized in a group dot plot. We can construct this based on code from AIBS

```{r, group dot plot}

devtools::install_github("AllenInstitute/scrattch.vis") # as needed
library(scrattch.vis)
options(stringsAsFactors = F) # following https://github.com/AllenInstitute/scrattch.vis

AIBS_lnCPM_InEx_ACC <- t(as.data.frame(Seu_AIBS_obj_ACC_it2[["RNA"]]@data)) #get transposed lnCPM matrix

library(tibble)
AIBS_lnCPM_InEx_ACC <- rownames_to_column(as.data.frame(AIBS_lnCPM_InEx_ACC))
colnames(AIBS_lnCPM_InEx_ACC)[1] <- "sample_name"
rownames(AIBS_lnCPM_InEx_ACC) <- AIBS_lnCPM_InEx_ACC$sample_name

gdp_anno <- tome_sample_meta_filtered
colnames(gdp_anno)[35] <- "query_subclass_label"
gdp_anno$query_subclass_id <- gdp_anno$query_subclass_label
gdp_anno$query_subclass_color <- "#C8C8C8"

group_dot_plot(AIBS_lnCPM_InEx_ACC, 
               gdp_anno, 
               genes = c("PVALB","SST"), 
               grouping = "query_subclass", 
               log_scale = TRUE,
               font_size = 10,
               max_size = 10,
               rotate_counts = TRUE)

```

This is the end of this notebook/our current workflow.