---
title: "Finding markers with SCMarker"
author: "Sonny"
output: html_notebook
---

This notebook/script places after "Loading_AIBS_Data.Rmd", and was written after "Seurat_Implementation.rmd", though it should be independent of the latter. We are now trying to find cell subclass markers with SCMarker and the AIBS data. 

### Installing the package

First, we install the SCMarker package, which is not trivial if working with R 3.6+ (as of the time of writing this document). 

For installtion in R 3.5 or earlier, the either of the following (untested by Sonny) two pairs of lines of code should work:

```{r, old R installation of SCMarker}

library(devtools)
install_github("KChen-lab/SCMarker")

install.packages("remotes")
remotes::install_github("KChen-lab/SCMarker")

```

For installation in R 3.6+, we need both bash and r code; we need to download and install the package rather than go through simple r installtion commands.

In bash, we do the following (code not actually run/tested for this document, since code was already run by the time of writing up this document).

We first allow bash script in this rmd.

```{r, lang_setup}

knitr::knit_engines
names(knitr::knit_engines$get())

```

Now in bash, we download the package in the working directory of this project.

```{bash, download package}

git clone https://github.com/KChen-lab/SCMarker.git

```

We can now build and install the package (code not actually run/tested for this document, since code was already run by the time of writing up this document).

```{r, SCMarker package build and installation}

library(devtools)
build("SCMarker")
#this should generate a .tar.gz file in your working/current directory. You can then install this in RStudio by graphic interface (packages tab, click install, etc.)

```

### Running the package/finding markers

We can now try using the package to find markers. Code developed/adjusted from https://github.com/KChen-lab/SCMarker. From their code, we seem to simply need to input a normalized count matrix, formatted as a matrix (and not a dataframe).

```{r, finding markers}

library(SCMarker)

res=ModalFilter(data= as.matrix(AIBS_Rawcount_InEx[1:10000,1:10000]) ,geneK=10,cellK=10,width=2) #from package creators: default width = 1 for UMI data, width =2 for TPM data.
res=GeneFilter(obj=res)
res=getMarker(obj=res,k=300,n=30)
head(res$marker)

```

